Rapid technological progress in white light-emitting diodes \\ and its sources in innovation and technology spillovers

### 2023-03-14 2_SSL_EES/article/article_supplementary.tex
added figure
updated figure

\documentclass[10pt]{article}
% page setup
\usepackage[a4paper, total={6in, 8.75in}]{geometry}
\usepackage{parskip}
% formatting
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc} % use 8-bit T1 fonts
% cross-referencing
\usepackage{hyperref}
\usepackage{url}
\usepackage{doi}
% tables
\usepackage{booktabs}
% fonts
\usepackage{amsfonts}
\usepackage{microtype}
% s
\usepackage{graphicx}
\usepackage{float}
% math
\usepackage{amsmath}
% tables
\usepackage{tabularx}
\usepackage{colortbl}
% boxes
\usepackage{fancybox}
% tikz
\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes,shapes.geometric,shapes.misc}
% landscape pages
\usepackage{pdflscape}
% chemical formulas
\usepackage{chemformula}
% smart references
\usepackage{cleveref}

\renewcommand{\figurename}{Figure SI}
\renewcommand{\tablename}{Table SI}
\renewcommand{\theequation}{SI \arabic{equation}}

@@

\begin{equation}
\label{eqn:ler}
    K [\text{lm/W}_{opt}]= \frac{\int_0^\infty K( \lambda ) \phi \text{d} \lambda}{\int_0^\infty \phi \text{d} \lambda}
\end{equation}

@@

\begin{equation}
    \eta [\text{lm/W}_{el}]= \frac{\phi}{P_{el}}
\end{equation}

@@

\begin{figure}[h!]
    \tikzfig{./figures/costmodel}
    \centering
    \caption{Schematic diagram of the cost model showing inputs to each step and computational steps leading to cost model outputs. For description of cost variables, see definitions for equation \cref{eqn:cost_wafer}.}
    \label{fig:costmodel-schematic}
    \vspace{-8pt}
\end{figure}

@@

@@

Figures \cref{fig:manuf_vtf_2012-1}-\cref{fig:manuf_csp_2020-2} show a simplified rendering of the manufacturing process of three different chip architectures considered in the cost model.

@@

\begin{equation}
\label{eqn:cost_sum}
    C \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = P_S + C_w + C_p
\end{equation}

@@

\begin{align}
	C_w &= \sum_i C_i \\
	C_p &= \sum_j C_j
\end{align}

@@

\begin{equation}
\label{eqn:cost_wafer}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] =\frac{1}{D}  \frac{1}{y_i}   \bigg\{ \bigg((e*p) + l + m + d +o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x p_x \bigg\}
\end{equation}

@@

\begin{equation}
	A_{\text{usable}}=A_{\text{wafer}}-A_{\text{cut}}-A_{\text{exclusion}}
\end{equation}

@@

\begin{equation}
\label{eqn:dpw}
	D=\frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2}
\end{equation}

@@

Results are plotted in Figure \cref{fig:dpw}, showing the increase in die per wafer (DPW). Equation \cref{eqn:dpw} then gives us for the cost of a manufacturing step $C_i$ in the wafer processing category

\begin{equation}
\label{eqn:cost_wafer_full}
\begin{split}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] &= \bigg (  \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2} \bigg )^{-1} \times \\
    &  \frac{1}{y_i}  \bigg\{ \bigg((e*p) + l + m + d +o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x p_x \bigg\}
\end{split}
\end{equation}

@@

\begin{equation}
\label{eqn:cost_die}
    C_j \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = \frac{1}{y_j}  \bigg\{ \bigg((e*p) + l + m + d + o \bigg)_i  \frac{c_j}{u_j} + \sum_{x} a v_x p_x \bigg\}
\end{equation}

where $c_j$ is $\text{throughput}^{-1}$. The total cost is thus
\begin{equation}
\label{eqn:cost_total}
\begin{split}
    C= P_s &+ \sum_i \bigg \{ \frac{1}{D} \frac{1}{y_i} \bigg[ \frac{t_i}{w_i u_i} \bigg((e*p) + l + m + d +o \bigg)_i  + \sum_{x} v_x p_x \bigg] \bigg \} + \\
    & + \sum_j \bigg \{ \frac{1}{y_j} \bigg[ \frac{c_j}{u_j}  \bigg((e*p) + l + m + d + o \bigg)_i + \sum_{x} a v_x p_x \bigg ] \bigg\}
\end{split}
\end{equation}

@@

\begin{equation}
\begin{split}
\label{eqn:C_2}
    C_{Y_1} &= \frac{C_1}{Y_1} \\
    C_{Y_2} &= \frac{C_1 + C_2}{Y_1 Y_2} - C_{Y_1} = \frac{C_1 + C_2}{Y_1 Y_2} - \frac{C_1}{Y_1} = \frac{1}{Y_1 Y_2} \bigg ( C_1 (1-Y_2) +C_2 \bigg)\\
    C_{Y_i} &= \frac{ \sum_{x \leq i} C_x }{ \prod_{x \leq i} Y_x } - \frac{ \sum_{x<i} C_x }{ \prod_{x<i} Y_x }
\end{split}
\end{equation}

@@

\begin{align}
\label{eqn:C_2^2}
    C_{Y_2}^{(2 \times)} &= \bigg( \frac{C_1 + C_2}{Y_1 Y_2} - \frac{C_1}{Y_1} \bigg) + \bigg( \frac{C_1 + 2 C_2}{Y_1 Y_2^2} - \frac{C_1 + C_2}{Y_1 Y_2}     \bigg) \\
    &= \frac{1}{Y_1 Y_2^2} \bigg( C_1 (1-Y_2^2) +2C_2 \bigg)
\end{align}

@@

\begin{equation}
    C_{Y_2}^{(n \times)} = \frac{1}{Y_1 Y_2^n} \bigg( C_1 (1-Y_2^n)+nC_2\bigg)
\end{equation}

@@

\begin{equation}
\label{eqn:yielded_cost}
    C_{Y_{i>1}}^{(n \times)} = \frac{1}{Y_i^{n-1} \prod_{x \leq i} Y_x} \bigg( nC_i + \sum_{x < i} C_x (1-Y_i^n) \bigg)
\end{equation}

@@

\begin{equation}
	f_\text{single} = \frac{i+s}{y}
\end{equation}

@@

\begin{align}
\label{eqn:series}
    f_{\times 2} &=  \frac{\frac{i+s}{y}+s}{y} \\
    f_{\times n} &= \frac{i + s(1+y+y^2+ \dots + y^{n-1})}{y^n}
\end{align}

@@

\begin{equation}
    f_\text{total} = \frac{\sum_i^n s_i}{\prod_i^ny_i} \neq \sum_i^n f_i
\end{equation}

@@

\begin{equation}
C=ab+cd
\end{equation}

@@

\begin{equation}
\Delta C = C(t_1)-C(t_0) = \sum_{i=a, \dots, d} \Delta C_i
\end{equation}

@@

\begin{equation}
\text{d}C(x_1 (t), x_2(t), \dots) = \sum_i \frac{\partial C }{\partial x_i}     \frac{\text{d}x_i}{\text{d}t} = \sum_i \frac{\partial C }{\partial x_i}  \Delta x_i
\end{equation}

@@

\begin{equation}
\Delta C_{x_1} = \int_{t=t_1}^{t_2} \frac{\partial C }{\partial x_1} \frac{\text{d}x_1}{\text{d}t} \text{d}t
\label{eqn:integral_1}
\end{equation}

@@

\begin{equation}
C(\vec{r}) = C(r_1,r_2, \dots) = \sum_i C_i
\end{equation}

@@

\begin{equation}
C_i(\vec{r}) = C_i^0 \prod_w g_{iw}(r_w)
\end{equation}

@@

\begin{equation}
\Delta C_x = \int_{t=t_0}^{t_1} C(t) \frac{ \partial \ln C }{ \partial x } \frac{ \text{d} x }{ \text{d} t} \text{d} t
\end{equation}

@@

\begin{equation}
\Delta C_z (t_1,t_2) \approx \sum_i \tilde{C_i} \ln \frac{g_{iz}(t_2)}{g_{iz}(t_1)}
\end{equation}

@@

Figure \cref{fig:costmodel_calibration} shows a comparison of the relative manufacturing cost by manufacturing step categories between the cost model and data released by the United States Department of Energy (DoE). Note that any comparison between the s must consider: Our model assumes state-of-the-art equipment, while the DoE industry round table reports data from factories which use older equipment, as well as newer equipment. Our model presently considers only low-power and mid-power packages, while the DoE reports consider high-power packages also. Our model assumes a virtual North American manufacturing location, while the majority of manufacturers produce in Asia. We note that the share of the substrate price in the DoE data is much larger than in our model. This can in part be explained by  \cref{fig:sapphire_prices}, which already shows how projections overestimated the actual price for 2015-2020. We note also that the share of the epitaxy step is larger in the DoE model. This can in part be explained by our model relying on state-of-the-art equipment, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. In conclusion, our model fits industry forecasts due to the way it is set up. The packaging part of the manufacturing process is increasing in importance, which has been confirmed by researchers and industry reports on WLP\cite{Lee2011WPL}\cite{Xie2013}\cite{ledsmag2017WLP}.

>>>>

\documentclass[10pt]{article}
% page setup
\usepackage[a4paper, total={6in, 8.75in}]{geometry}
\usepackage{parskip}
% formatting
\usepackage[utf8]{inputenc} % allow utf-8 input
\usepackage[T1]{fontenc} % use 8-bit T1 fonts
% cross-referencing
\usepackage{hyperref}
\usepackage{url}
\usepackage{doi}
% tables
\usepackage{booktabs}
% fonts
\usepackage{amsfonts}
\usepackage{microtype}
% s
\usepackage{graphicx}
\usepackage{float}
% math
\usepackage{amsmath}
% tables
\usepackage{tabularx}
\usepackage{colortbl}
% boxes
\usepackage{fancybox}
% tikz
\usepackage{tikz}
\usetikzlibrary{backgrounds}
\usetikzlibrary{arrows}
\usetikzlibrary{shapes,shapes.geometric,shapes.misc}
% landscape pages
\usepackage{pdflscape}
% chemical formulas
\usepackage{chemformula}
% smart references
\usepackage[capitalise, nameinlink]{cleveref}

\renewcommand{\figurename}{Figure SI}
\renewcommand{\tablename}{Table SI}

% custom equation numbering
% https://tex.stackexchange.com/a/184641/
\newcounter{defcounter}
\setcounter{defcounter}{0}
\newenvironment{myequation}{%
\addtocounter{equation}{-1}
\refstepcounter{defcounter}
\renewcommand\theequation{SI\thedefcounter}
\begin{equation}}
{\end{equation}}

@@

\begin{myequation}
\label{eqn:ler}
    K [\text{lm/W}_{opt}]= \frac{\int_0^\infty K( \lambda ) \phi \text{d} \lambda}{\int_0^\infty \phi \text{d} \lambda}
\end{myequation}

@@

\begin{myequation}
    \eta [\text{lm/W}_{el}]= \frac{\phi}{P_{el}}
\end{myequation}

@@

\begin{figure}[h!]
    \centering
    \includegraphics[width=0.95\textwidth]{./figures/cost_model.pdf}
    \caption{Schematic diagram of the cost model showing inputs to each step and computational steps leading to the cost model outputs. For the description of cost variables, see definitions for equation \cref{eqn:cost_wafer}.}
    \label{fig:costmodel-schematic}
\end{figure}

@@

Calling something here as 'upstream' and 'downstream' is kind of confusing. I suggest to replace the text highlighted by this comment with the following simplifying reformulation: 
'It is split between the two stages of the manufacturing process: first stage combining operations at the wafer level, followed by the LED packaging stage.'

@@

\cref{fig:manuf_vtf_2012-1}-\cref{fig:manuf_csp_2020-2} show a simplified rendering of the manufacturing process of three different chip architectures considered in the cost model.

    %CLASSICAL CHIP

    \begin{landscape}
        \begin{figure}
            \includegraphics[width=555pt]{./figures/classical_overview_2003.pdf}
            \caption{Manufacturing process for a classical LED package with lateral current spreading, circa 2003. Abbreviations: MQW - multiple quantum well; PLLC - plastic leaded chip carrier. Continued on next page.}
            \label{fig:manuf_classical_2003}
        \end{figure}
    \end{landscape}

@@

\begin{myequation}
\label{eqn:cost_sum}
    C \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = P_S + C_w + C_p
\end{myequation}

@@

\begin{myequation}
        C_w = \sum_i C_i \\
	C_p = \sum_j C_j
\end{myequation}

@@

\begin{myequation}
\label{eqn:cost_wafer}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] =\frac{1}{D}  \frac{1}{y_i}   \bigg\{ \bigg((e*p) + l + m + d +o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x p_x \bigg\}
\end{myequation}

@@

\begin{myequation}
	A_{\text{usable}}=A_{\text{wafer}}-A_{\text{cut}}-A_{\text{exclusion}}
\end{myequation}

@@

\begin{myequation}
\label{eqn:dpw}
	D=\frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2}
\end{myequation}

@@

Results are plotted in \cref{fig:dpw}, showing the increase in die per wafer (DPW). \cref{eqn:dpw} then gives us for the cost of a manufacturing step $C_i$ in the wafer processing category

\begin{myequation}
\label{eqn:cost_wafer_full}
\begin{split}
    C_i \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] &= \bigg (  \frac{\pi}{4}  \bigg ( \frac{d-2e}{\sqrt{a}+s/2} \bigg ) ^2 - \frac{\pi}{\sqrt{2}}\frac{d-2e}{(\sqrt{a}+s/2)^2} \bigg )^{-1} \times \\
    &  \frac{1}{y_i}  \bigg\{ \bigg((e*p) + l + m + d +o \bigg)_i \bigg( \frac{t_i}{w_i u_i} \bigg) + \sum_{x} v_x p_x \bigg\}
\end{split}
\end{myequation}

@@

\begin{myequation}
\label{eqn:cost_die}
    C_j \bigg[ \frac{ \text{USD}(2020) }{ \text{die} } \bigg] = \frac{1}{y_j}  \bigg\{ \bigg((e*p) + l + m + d + o \bigg)_i  \frac{c_j}{u_j} + \sum_{x} a v_x p_x \bigg\}
\end{myequation}

where $c_j$ is $\text{throughput}^{-1}$. The total cost is thus
\begin{myequation}
\label{eqn:cost_total}
\begin{split}
    C= P_s &+ \sum_i \bigg \{ \frac{1}{D} \frac{1}{y_i} \bigg[ \frac{t_i}{w_i u_i} \bigg((e*p) + l + m + d +o \bigg)_i  + \sum_{x} v_x p_x \bigg] \bigg \} + \\
    & + \sum_j \bigg \{ \frac{1}{y_j} \bigg[ \frac{c_j}{u_j}  \bigg((e*p) + l + m + d + o \bigg)_i + \sum_{x} a v_x p_x \bigg ] \bigg\}
\end{split}
\end{myequation}

@@

\begin{myequation}
\label{eqn:C_2}
    C_{Y_1} = \frac{C_1}{Y_1}
\end{myequation}
\begin{myequation}
    C_{Y_2} = \frac{C_1 + C_2}{Y_1 Y_2} - C_{Y_1} = \frac{C_1 + C_2}{Y_1 Y_2} - \frac{C_1}{Y_1} = \frac{1}{Y_1 Y_2} \bigg ( C_1 (1-Y_2) +C_2 \bigg)
\end{myequation}
\begin{myequation}
    C_{Y_i} = \frac{ \sum_{x \leq i} C_x }{ \prod_{x \leq i} Y_x } - \frac{ \sum_{x<i} C_x }{ \prod_{x<i} Y_x }
\end{myequation}

@@

\begin{myequation}
\label{eqn:C_2^2}
    C_{Y_2}^{(2 \times)} = \bigg( \frac{C_1 + C_2}{Y_1 Y_2} - \frac{C_1}{Y_1} \bigg) + \bigg( \frac{C_1 + 2 C_2}{Y_1 Y_2^2} - \frac{C_1 + C_2}{Y_1 Y_2}     \bigg)
\end{myequation}
\begin{myequation}
    = \frac{1}{Y_1 Y_2^2} \bigg( C_1 (1-Y_2^2) +2C_2 \bigg)
\end{myequation}

@@

\begin{myequation}
    C_{Y_2}^{(n \times)} = \frac{1}{Y_1 Y_2^n} \bigg( C_1 (1-Y_2^n)+nC_2\bigg)
\end{myequation}

@@

\begin{myequation}
\label{eqn:yielded_cost}
    C_{Y_{i>1}}^{(n \times)} = \frac{1}{Y_i^{n-1} \prod_{x \leq i} Y_x} \bigg( nC_i + \sum_{x < i} C_x (1-Y_i^n) \bigg)
\end{myequation}

@@

\begin{myequation}
	f_\text{single} = \frac{i+s}{y}
\end{myequation}

@@

\begin{myequation}
\label{eqn:series}
    f_{\times 2} =  \frac{\frac{i+s}{y}+s}{y} \\
    f_{\times n} = \frac{i + s(1+y+y^2+ \dots + y^{n-1})}{y^n}
\end{myequation}

@@

\begin{myequation}
    f_\text{total} = \frac{\sum_i^n s_i}{\prod_i^ny_i} \neq \sum_i^n f_i
\end{myequation}

@@

\begin{myequation}
C=ab+cd
\end{myequation}

@@

\begin{myequation}
\Delta C = C(t_1)-C(t_0) = \sum_{i=a, \dots, d} \Delta C_i
\end{myequation}

@@

\begin{myequation}
\text{d}C(x_1 (t), x_2(t), \dots) = \sum_i \frac{\partial C }{\partial x_i}     \frac{\text{d}x_i}{\text{d}t} = \sum_i \frac{\partial C }{\partial x_i}  \Delta x_i
\end{myequation}

@@

\begin{myequation}
\Delta C_{x_1} = \int_{t=t_1}^{t_2} \frac{\partial C }{\partial x_1} \frac{\text{d}x_1}{\text{d}t} \text{d}t
\label{eqn:integral_1}
\end{myequation}

@@

\begin{myequation}
C(\vec{r}) = C(r_1,r_2, \dots) = \sum_i C_i
\end{myequation}

@@

\begin{myequation}
C_i(\vec{r}) = C_i^0 \prod_w g_{iw}(r_w)
\end{myequation}

@@

\begin{myequation}
\Delta C_x = \int_{t=t_0}^{t_1} C(t) \frac{ \partial \ln C }{ \partial x } \frac{ \text{d} x }{ \text{d} t} \text{d} t
\end{myequation}

@@

\begin{myequation}
\Delta C_z (t_1,t_2) \approx \sum_i \tilde{C_i} \ln \frac{g_{iz}(t_2)}{g_{iz}(t_1)}
\end{myequation}

@@

\cref{fig:costmodel_calibration} shows a comparison of the relative manufacturing cost by manufacturing step categories between the cost model and data released by the United States Department of Energy (DoE). Note that any comparison between the s must consider: Our model assumes state-of-the-art equipment, while the DoE industry round table reports data from factories which use older equipment, as well as newer equipment. Our model presently considers only low-power and mid-power packages, while the DoE reports consider high-power packages also. Our model assumes a virtual North American manufacturing location, while the majority of manufacturers produce in Asia. We note that the share of the substrate price in the DoE data is much larger than in our model. This can in part be explained by  \cref{fig:sapphire_prices}, which already shows how projections overestimated the actual price for 2015-2020. We note also that the share of the epitaxy step is larger in the DoE model. This can in part be explained by our model relying on state-of-the-art equipment, while industry might not run low-power and mid-power chip production on these, more expensive, reactors. In conclusion, our model fits industry forecasts due to the way it is set up. The packaging part of the manufacturing process is increasing in importance, which has been confirmed by researchers and industry reports on WLP\cite{Lee2011WPL}\cite{Xie2013}\cite{ledsmag2017WLP}.
